---
title: "CI coverage simulation"
author: "Dr. Matthew Beckman"
output: html_notebook
---


# Simulation 

#### Result from a real study about carbon capture in tropical forests: 

> 95% CI: 11,600 ± 2⋅1000 = (9,600, 13,600)

> "We are 95% confident that the average amount of carbon stored in each square kilometer of tropical forest is between 9,600 and 13,600 tons"


What's wrong with saying "there's a 95% chance that the average amount of carbon stored in each square kilometer of tropical forest is between 9,600 and 13,600 tons" ??

```{r}
normDatSim <- rnorm(n = 15, mean = 11600, sd = 1000)
# normDatSim <- rexp(n = 15, rate = 1/11600)  # a non-normal distribution for comparison

ggplot() +
  geom_density(aes(normDatSim))

t.test(normDatSim)

```


# Simulation with Loops

```{r}
# simulation settings

nsimulations <- 100
mu <- 11600
sigma <- 1000
sampleSize <- 15

# initialize a data frame where we can store various results at each step in our simulation

ConfIntDat <- data.frame(simulationID = rep(NA, nsimulations), 
                         lower = rep(NA, nsimulations), 
                         upper = rep(NA, nsimulations), 
                         captureMu = rep(NA, nsimulations))

# use a loop to simulate many results and store them at each step

for (i in 1:nsimulations) {                                   # increment 'i' from 1 to 'nsimulations'
  normDatSim <- rnorm(n = sampleSize, mean = mu, sd = sigma)  # simulate a new random sample
  ConfIntDat$simulationID[i] <- i                             # store an ID for each sample
  ConfIntDat$lower[i] <- t.test(normDatSim)$conf.int[1]       # store the lower bound of our CI
  ConfIntDat$upper[i] <- t.test(normDatSim)$conf.int[2]       # store upper bound of our CI
  ConfIntDat$captureMu[i] <- (mu > ConfIntDat$lower[i] & mu < ConfIntDat$upper[i])  # is mu captured?

}

ConfIntDat

# let's plot it!

library(tidyverse)

ConfIntDat %>%
  ggplot() +
  geom_segment(aes(    x = lower,    y = simulationID, 
                    xend = upper, yend = simulationID, 
                   color = captureMu)) + 
  geom_vline(aes(xintercept = mu)) + 
  xlab("Simulated confidence interval from Normal data") + 
  ylab("Index of a CI based on a simulated random sample")


# how often did our 95% CI actually capture mu? Easy!

mean(ConfIntDat$captureMu)


```

- Can you see why we call it a "95% confidence interval"?
- Tinker with the simulation settings and try again 
    - Explore changes to `nsimulations`, `mu`, `sigma`, and `sampleSize` 
    - Try changing one at a time to larger or smaller values and observe what happens


# another way... (some would say "The R Way")

- we can make a function that handles the guts of the loop and then call the function as an argument 
- then we can leverage more efficient tools to do things way faster (vectorized functions, parallel processing, etc)



### making a function...

- the simulation settings will be arguments we (or another user) might want to change
- when in doubt, it's better to program arguments rather than "hard code" specific values within the guts of the function

```{r}

ciSim <- function(mu, sigma, sampleSize) {
  ## Purpose: function simulates random Normal data and constructs a 95% CI
  ## Arguments: 
  ##    mu: population mean for Normal distribution from which samples are drawn
  ##    sigma: population standard deviation for Normal distribution from which samples are drawn
  ##    sampleSize: sample size to draw from Normal distribution
  
  result <- list(NULL)
  normDatSim <- rnorm(n = sampleSize, mean = mu, sd = sigma)  # simulate a new random sample
  result[1] <- t.test(normDatSim)$conf.int[1]                 # store lower bound of CI
  result[2] <- t.test(normDatSim)$conf.int[2]                 # store upper bound of CI
  result[3] <- (mu > result[1] & mu < result[2])              # store mu capture result
  return(result)                                              # return our results
}

ciSim(mu = 34, sigma = 5, sampleSize = 15)

```

### let's use mosaic::do() which can "do" something many times without writing a loop

```{r}
library(mosaic)

sim_results <- mosaic::do(10000) * ciSim(mu = 34, sigma = 5, sampleSize = 15)
sim_results

# some handy tricks to get things into a data frame as desired
sim_results_df <- 
  sim_results %>%
  transmute(lower = unlist(V1),    # unlist() simplifies to the object class inside the list element
            upper = unlist(V2), 
            muCapture = unlist(V3)) %>%
  rownames_to_column() %>%    # handy function to create a column that indexes each row
  mutate(rowname = parse_number(rowname))


sim_results_df

```



### plot it


```{r}
sim_results_df %>%
  ggplot() +
  geom_segment(aes(    x = lower,    y = rowname, 
                    xend = upper, yend = rowname, 
                   color = muCapture)) + 
  geom_vline(aes(xintercept = 34)) +      # mu reference is hard coded here because mu wasn't returned by `ciSim()`
  xlab("Simulated confidence interval from Normal data") + 
  ylab("Index of a CI based on a simulated random sample")


# how often did our 95% CI actually capture mu? Easy!

mean(sim_results_df$muCapture)

```





