---
title: "Week 8: Back to `ggplot2`"
subtitle: "Data Computing Chapter 8"
author: |
        | Presenter: Omar Hagrass
        | Content credit: Matthew Beckman
output: 
    slidy_presentation: default
    html_notebook: default
---


```{r include=FALSE}
library(tidyverse)
library(dcData)
library(mosaic)
library(esquisse)
knitr::opts_chunk$set(tidy=FALSE, message=FALSE)
options(width = 80)
```

----


## Agenda

- `ggplot()` flyover
- remarks about facets


## Building Graphics

1. Draw by hand (or imagine) the specific plot that you intend to construct
2. Data Wrangling (if needed) to get the data in glyph-ready form, or verify that the current form is glyph-ready for your purposes.
3. Establish the frame using a `ggplot()` statement
4. Create the intended glyph using `geom_[style]()` such as
    - `geom_point()`
    - `geom_bar()`
    - `geom_boxplot()`
    - `geom_density()`
    - `geom_vline()`
    - `geom_segment()`
    - `geom_histogram()`
    - and *many* more
5. Map variables to the graphical attributes of the glyph using: `aes( )`
6. Add additional layers to the frame using the `+` symbol 
    - Note: **not** `%>%` between layers of `ggplot2` graphics
    - Maybe think "add layer" in `ggplot2` portions, instead of "and then" with `%>%` syntax


## Want some help?

This can all be a lot to keep straight.  Best to learn by doing, but take advantage of available resources to help shoulder the burden:

- [EnoughDCF](http://dtkaplan.github.io/DCF-2014/CourseNotes/Guides/EnoughDCF.pdf): one page summary designed to accompany the Data Computing book.  Visit "Canvas >> Resources" for access.

- [RStudio Cheat Sheets](https://www.rstudio.com/resources/cheatsheets/), especially:
    - Data Visualization: 2 page summary of `ggplot2` tools and usage from RStudio
    - Data Transformation: 2 page summary of `dplyr` (& `tidyr`) tools and usage from RStudio

Also see...  

- [rstudio-rmarkdown-cheatsheet](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf)
- [rstudio-rmarkdown-reference](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)
- [rstudio-IDE-cheatsheet](https://www.rstudio.com/wp-content/uploads/2016/01/rstudio-IDE-cheatsheet.pdf)
- [and more...](https://www.rstudio.com/resources/cheatsheets/)




## Remarks about faceting: `facet_wrap()`

The syntax for facets requires a formula syntax we haven't seen much yet.  Also, there are two main ways to plot with facets.  Here are a few pointers:

- `facet_wrap()` just makes a box for each level of the categorical variable
    - Syntax: `facet_wrap( ~ categoricalVariable)`
    - For example:

```{r}
data("NCHS")

# 1is.na(smoker) gets cases that are non-missing for `smoker` (i.e. removes NA's)
Heights <- 
  NCHS %>%
  filter(age > 20, !is.na(smoker)) %>%   
  group_by(sex, smoker, age) %>%
  summarise(height = mean(height, na.rm = TRUE))

Heights %>%
  ggplot(aes(x = age, y = height)) +   
  geom_line(aes(linetype = smoker)) +   
  facet_wrap( ~ sex)
```

## Remarks about faceting: `facet_grid()`

- `facet_grid()` allows control of row & column facets
- `facet_grid()` syntax:
    - row & column facets: `facet_grid(rows ~ cols)`
    - row facets only: `facet_grid( rows ~ . )` (note the required "`.`") 
    - column facets only: `facet_grid( ~ cols)` (no "`.`" this time)

```{r}
Heights %>%
  ggplot(aes(x = age, y = height)) + 
  geom_line(aes(linetype = smoker)) + 
  facet_grid(sex ~ .)
```

```{r}
Heights %>%
  ggplot(aes(x = age, y = height)) + 
  geom_line() + 
  facet_grid(sex ~ smoker)
```


## Back to `BabyNames`

[Live example in class]



## Sky's the limit


Once you've mastered `ggplot2` you'll have incredible control to make all sorts of interesting plots!

- Fun stuff: <https://github.com/chang/ggphysics>
- Real article about the `BabyNames` data in popular media (`ggplot2` graphics):

<https://fivethirtyeight.com/features/how-to-tell-someones-age-when-all-you-know-is-her-name/>







