---
title: "Week 8 Class Notes"
output:
  html_document:
    fig_caption: yes
    theme: flatly
  pdf_document: default
---

```{r include=FALSE}
# Frontmatter
require(mosaic)
require(knitr)
require(dplyr)
require(tidyr)

# For typesetting
library(xtable)
library(ggthemes)
baseTheme <- theme_tufte( ticks=TRUE )
library(xkcd)
funTheme <- theme_xkcd()
library(scorer)
# opts_chunk$set( tidy=FALSE, out.width="50%", comment=NA,
#                warning=FALSE)
options( width=60, 
         na.rm=TRUE,
         # Printing tables
         xtable.type="html",
         xtable.include.rownames=FALSE)
require(DataComputing)
```

## Announcement

- Office hours update:
    - Mondays 11am - noon (316 Thomas)
    - Tuesdays 3pm - **4pm** (330 Thomas)
    - Wednesdays 2:30pm - **4pm** (316 Thomas)
    - **Thursdays 3:30pm - 4:30pm (330 Thomas)**
    - Fridays 11am - noon (316 Thomas)
- Feel free to use office hours as a chance to just meet with your group and work together with an expert nearby if needed. 
- Homework Solutions, Activity Examples, & R Cheat Sheets: "Canvas >> Files"


## Building Graphics

1. Draw (by hand) or imagine the specific plot that you intend to construct
#. Data Wrangling (if needed) - get the data in glyph-ready form, or verify that the current form is glyph-ready for your purposes.
#. Establish the frame using a `ggplot()` statement
#. Create the intended glyph using `geom_[style]()` such as
    * `geom_point()`
    * `geom_bar()`
    * `geom_boxplot()`
    * `geom_density()`
    * `geom_vline()`
    * `geom_segment()`
    * `geom_histogram()`
    * `geom_ash()`
    * and *many* more
#. Map variables to the graphical attributes of the glyph using `aes( )`
#. Add additional layers to the frame using the `+` symbol


## Want some help?

This can all be a lot to keep straight... visit "Canvas >> Files >> Cheatsheets"" for some resources to help you shoulder the burden:

- [EnoughDCF](http://dtkaplan.github.io/DCF-2014/CourseNotes/Guides/EnoughDCF.pdf): one page summary from the author of Data Computing
- [rstudio-ggplot2-cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/12/ggplot2-cheatsheet-2.0.pdf): 2 page summary of `ggplot2` tools and usage from RStudio
- [rstudio-data-wrangling-cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf): 2 page summary of `dplyr` (& `tidyr`) tools and usage from RStudio

Also see...  

- [rstudio-rmarkdown-cheatsheet](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf)
- [rstudio-rmarkdown-reference](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)
- [rstudio-IDE-cheatsheet](https://www.rstudio.com/wp-content/uploads/2016/01/rstudio-IDE-cheatsheet.pdf)
- [and more...](https://www.rstudio.com/resources/cheatsheets/)



## Remarks about `facet_wrap()`

The syntax for facets can be a bit complicated for new users.  Here are a few pointers:

- `facet_wrap()` just makes a box for each level of the categorical variable
    - Syntax: `facet_wrap( ~ categoricalVariable)`
    - For example:

```{r}
data("NCHS")

# 1is.na(smoker) gets cases that are non-missing for `smoker` (i.e. removes NA's)
Heights <- 
  NCHS %>%
  filter(age > 20, !is.na(smoker)) %>%   
  group_by(sex, smoker, age) %>%
  summarise(height = mean(height, na.rm = TRUE))

Heights %>%
  ggplot(aes(x = age, y = height)) + 
  geom_line(aes(linetype = smoker)) + 
  facet_wrap( ~ sex)
```

## Remarks about `facet_grid()`

- `facet_grid()` allows control of row & column facets
- `facet_grid()` syntax:
    - row & column facets: `facet_grid(rows ~ cols)`
    - row facets only: `facet_grid( rows ~ . )` (note the required "`.`") 
    - column facets only: `facet_grid( ~ cols)` (no "`.`" this time)

```{r}
Heights %>%
  ggplot(aes(x = age, y = height)) + 
  geom_line(aes(linetype = smoker)) + 
  facet_grid(sex ~ .)
```

```{r}
Heights %>%
  ggplot(aes(x = age, y = height)) + 
  geom_line() + 
  facet_grid(sex ~ smoker)
```


## Homework

- Activity: Bicycle Sharing (DC p. 182)
    - Sections A.3 & A.5 are worth extra credit (not required)
    - Make sure you complete all the "your turn" segments in A.1, A.2, A.4
    - The last "your turn (optional)" at the end of A.2 is **required**
- **Homework**:  
    - DC Ch 8 Exercises: 8.1, 8.2, 8.3 (skip the world map)
    - DC chapter 11 reading quiz on Canvas    

## Activity: Bicycle Sharing (DC p. 182)

Assignment is worth a total of 10 points.

* [2 point] Turn in HTML with embedded .Rmd file (e.g. "DataComputing simple" template)
* [2 points] A.1 side-by-side boxplots of rental durations
* [4 points total] A.2 plots
    - [0.5 point] density plot: day of the year (1 to 365)
    - [0.5 point] density plot: day of the week (Sunday to Saturday; numeric is fine)
    - [0.5 point] density plot: hour of the day (0 to 24)
    - [0.5 point] density plot: minute in the hour (0 to 60)
    - [1 points] Plot rental frequency by hour, weekday, & client type
    - [1 points] Plot rental density by hour, weekday, & client type
* [2 points] A.4 leaflet
* [**EXTRA CREDIT**: 2 pts] complete both A.3 & A.5

* General Remarks: 
    - Sections A.3 & A.5 are worth extra credit (not required)
    - Make sure you complete all the "your turn" segments in A.1, A.2, A.4
    - The last "your turn (optional)" at the end of A.2 is **REQUIRED**
    

---

[teaching](https://mdbeckman.github.io/Teaching.html) | [stat 184 home](https://mdbeckman.github.io/Teaching/AY2016-17/stat184/stat184-fall2016.html) | [syllabus](https://mdbeckman.github.io/Teaching/AY2016-17/stat184/stat184-syllabus.html) | [piazza](https://piazza.com/psu/fall2016/stat184/home) | [canvas](https://psu.instructure.com/courses/1792768)

