---
title: "Worked Example In Class"
output: html_notebook
---



```{r}
# clean up 
rm(list = ls())

# load necessary packages
library(tidyverse)  # includes ggplot2
library(dcData)
library(mosaic)

# data intake
data("BabyNames", package = "dcData")


# inspect data intake
head(BabyNames)
glimpse(BabyNames)
```


```{r}
# wrangle into glyph-ready form
BeckmanNames <- 
  BabyNames %>%
  filter(name %in% c("Matthew", "Sarah", "Eden", "Jack", "Hazel")) %>%  
  group_by(name, year) %>%
  summarise(total = sum(count, na.rm = TRUE))

# inspect the result
BeckmanNames %>%
  head(6)
```


```{r}
# in the beginning you might use mplot to get started--here's the default result
ggplot(data=BeckmanNames,aes(x=total,y=year))+geom_point() 

```


```{r}
# we can do better
    # establish the frame
    # pick a glyph/geom
    # map aesthetics
BeckmanNames %>%
  ggplot() + 
  geom_line(aes(x = year, y = total, color = name))


```

## Lets overlay the overall birth trend

First we need to calculate the total number of US births each year--at least those with Social Security records.

```{r}
AnnualBirths <- 
  BabyNames %>%
  group_by(year) %>%
  summarise(annualTotal = sum(count, na.rm = TRUE))

```

Lets plot that trend...

```{r}
AnnualBirths %>%
  ggplot() +
  geom_line(aes(x = year, y = annualTotal))
```


Can we add that information as a context layer for the plot of `BeckmanNames`? 


```{r}
BeckmanNames %>%
  ggplot() + 
  geom_line(aes(x = year, y = total, color = name)) + 
  geom_line(data = AnnualBirths, aes(x = year, y = annualTotal / 50), alpha = 0.2) + 
  ggtitle(label = "Frequency of Names in the Beckman Home", 
          subtitle = "Including overlay of scaled (50:1) frequency of births in US (per SSA)")



```






## [From Slides] Sky's the limit!!

Once you've mastered `ggplot2` you'll have incredible control to make all sorts of interesting plots!

- Fun stuff: <https://github.com/chang/ggphysics>
- Real article about the `BabyNames` data in popular media (`ggplot2` graphics):

<https://fivethirtyeight.com/features/how-to-tell-someones-age-when-all-you-know-is-her-name/>




## Here's a crack at that one

With support from Modern Data Science with R by Baumer, Horton, & Kaplan.

```{r}
library(babynames)  # includes slightly different version of babynames data
library(Hmisc)      # has a quantile function I want to use
library(mdsr)
library(ggthemes)   

BabynamesDist <- make_babynames_dist()  # from mdsr; take a look!
glimpse(BabynamesDist)

# Age distribution for American males named "Joseph" (according to US Social Security data)

Joseph <- BabynamesDist %>%
  filter(name == "Joseph" & sex == "M")

# constructing the plot
name_plot <- 
  Joseph %>%
  ggplot(aes(x = year)) + 
  geom_bar(stat = "identity", aes(y = count_thousands * alive_prob), 
           fill = "#b2d7e9", color = "white") + 
  geom_line(aes(y = count_thousands), size = 2) + 
  ylab("Number of People (thousands)") + 
  xlab(NULL)

# first pass
name_plot
```


```{r}
# we use the `wtd.quantile()` function from the `Hmisc` package here
median_yob <- wtd.quantile(x = Joseph$year, weights = Joseph$est_alive_today, probs = 0.5)
median_yob
```


```{r}
# note use of `ifelse` to create a new histogram with one non-zero bar at `median_yob`
# add titles
# add text labels (improve fonts)
# add curved arrow

name_plot <- 
  name_plot + 
  geom_bar(stat = "identity", color = "white", fill = "#008fd5", 
           aes(y = ifelse(year == median_yob, est_alive_today / 1000, 0))) 

name_plot + 
  ggtitle(label = "Age Distribution of American Boys Named Joseph", subtitle = "By year of birth") + 
  geom_text(x = 1935, y = 40, size = 3.5, family = "mono", label = "Number of Josephs \n born each year") + 
  geom_text(x = 1915, y = 13, size = 3.5, family = "mono", color = "#008fd5", 
            label = "Number of Josephs \n born each year \n estimated to be alive \n on Jan. 1, 2014") + 
  geom_text(x = 2003, y = 40, size = 3.5, family = "sans", color = "darkgray",
            label = "The median\nliving Joseph\nis 37 years old.") + 
  geom_curve(x = 1995, xend = 1974, y = 40, yend = 24, 
             arrow = arrow(length = unit(0.3, "cm")), curvature = 0.5) + 
  ylim(0, 42) + 
  theme_fivethirtyeight()  # getting pretty close!
  
```




