---
title: "Regular Expressions"
subtitle: "Data Computing Ch 17"
author: |
        | Presenter: Omar Hagrass
        | Content credit: Matthew Beckman
output: 
    slidy_presentation: default
    html_notebook: default
---


```{r include=FALSE}
require(mosaic)
require(lubridate)
require(tidyverse)
require(rvest)
knitr::opts_chunk$set(message = FALSE)
```


## Pole Vault Records Clean Up 

```{r echo=TRUE}
webpage <- "https://en.wikipedia.org/wiki/Men%27s_pole_vault_world_record_progression"

table_list <- 
  webpage %>%
  read_html(header=TRUE) %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)

PVRecords <- table_list[[1]]  # convert list to data frame
head(PVRecords, 3) # inspect the data

```

- we should fix the variable name representing the number of world records achieved by each athlete
- locate and replace all footnotes in the `Date` variable using `gsub()`
- convert `Date` to a date class variable in R using a `lubridate` function
- use `tidyr::extract()` to store the metric heights from the `Record` variable


```{r}

# locate and replace all footnotes in the `Date` column
PVMen <- 
  PVRecords %>%
  rename(recordsBroken = `#[4]`) %>%
  mutate(Date = gsub(pattern = "\\[.\\]", replacement = "", x = Date)) %>%
  mutate(Date = lubridate::mdy(Date)) %>% #convert to date
  tidyr::extract(col = Mark, into = "Meters", regex = "(^\\d\\.\\d)") %>%
  mutate(Meters = parse_number(Meters)) #convert to numeric(drops non-numeric characters)
  
PVMen %>% 
  head()
```





```{r}
PVMen %>%
  ggplot(aes(x = Date, y = Meters)) + 
  geom_step() + 
  geom_point(alpha = 0.5, aes(color = Nation))+
  scale_color_brewer(palette = "Spectral")

```

## What happened?


Here's a video explanation if you want to watch the evolution: 

<https://youtu.be/QGTdEhUW0nE>

