---
title: "Week 10: Web Scraping Demo"
subtitle: "STAT 184"
author: "Dr. Matthew Beckman"
output: 
    html_notebook: default
---

```{r include=FALSE}

knitr::opts_chunk$set(tidy=FALSE, message=FALSE, warning=FALSE)
options(width = 80)
```




## Penn State Football Receiving Statistics

1. Google Penn State Football Statistics: <http://bfy.tw/88gl>

2. Edit the R code chunk shown to paste `webpage` URL with quotes around it as shown.

3. Execute the code chunk to scrape all HTML tables found on the page into a "list" object in the R environment called `Tables` here

4. Identify a data table from the source (for example, "receiving statistics") and find it in the list object in your R environment


```
webpage <- ""

table_list <- 
  webpage %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)

str(table_list)
```


## Penn State Football Receiving Statistics


```{r}
library(tidyverse)
library(rvest)




```


```{r}
# R stores the result as a "list" object, let's take a look


# Double square brackets select an element of the list; let's store it at as a data frame


# Inspect result


# Tidy up the data table & rename variables  

  
# Inspect result


```

